Hybrid Partial Least Squares Regression with Multiple Functional and Scalar Predictors

### 2025-08-19 manuscript/main.tex
hilbert space proof

	\subsection{Iterative steps}
	The iterative process presented here yields an orthonormal hybrid basis that effectively captures the predictor-response relationships. It proceeds through two intermediate steps: the estimation of the PLS component direction (Section \ref{section:sub:compute_PLS_component}) and residualization (Section \ref{section:sub:residualization}). The properties of the resulting estimates are introduced in Section \ref{section:main:properties_of_our_method}.
	\subsubsection{Iterative step 1: regularized estimation of  PLS component direction}  \label{section:sub:compute_PLS_component} 
	We begin by formally introducing the core optimization problem pertinent to the PLS direction estimation, which is formulated as a generalized Rayleigh quotient (Proposition \ref{proposition:eigen_noregul}). Building upon this foundational concept, we present our regularized PLS component direction estimation step that promotes smoothness (Proposition \ref{proposition:eigen_regul}). Furthermore,  we detail an efficient computational scheme (Proposition \ref{proposition:linear_regul}).

@@

	\subsection{Final step: estimating  the hybrid regression coefficient}
	\label{section:sub:regression_coeff}
	The hybrid regression coefficient $\beta$ in model \eqref{eq: Hybrid functional model} can be written in terms of the estimated PLS directions and scores. First, note that we can show $
	\widehat{\rho}_i^{[l]}
	=
	\langle  W_i^{[l]}, \widehat{\xi}^{[l]} \rangle_\mathbb{H}
	=
	\langle  W_i, \widehat{ \iota}^{[l]} \rangle_\mathbb{H}
	$, where 
	$\widehat{ \iota}^{[l]} = \widehat{\xi}^{[l]} - \sum_{u=1}^{l-1} \langle \widehat{ \delta}^{[u]}, \widehat{\xi}^{[l]} \rangle_{\mathbb{H}} \widehat{ \iota}^{[u]}$ for $l \ge 2$, with $
	\widehat{\iota}^{[1]} = \widehat{\xi}^{[1]}
	$. Next,  \eqref{algorithm_step:residualization_predictor} leads to the following model:
	\begin{equation*}
		Y_i = \sum \limits_{l=1}^L \widehat{\nu}^{[l]} \widehat{\rho}_i^{[l]} + \epsilon_i.
	\end{equation*}
	This model lets us to express $Y_i$ as:
	\begin{equation*}
		Y_i = \sum \limits_{l=1}^L \widehat{\nu}^{[l]} \langle W_i^{[l]}, \widehat{\xi}^{[l]} \rangle_\mathbb{H} +\epsilon_i 
		= 
		\bigl \langle W_i, \sum \limits_{l=1}^L \widehat{\nu}^{[l]} \widehat{ \iota }^{[l]}
		\bigr \rangle_\mathbb{H}+\epsilon_i,
	\end{equation*}
	which, given the uniqueness of $ \beta $, leads to
	\begin{equation*}
		\widehat{ \beta} = \sum \limits_{l=1}^L\widehat{\nu}^{[l]} \widehat{ \iota }^{[l]}.
	\end{equation*}
	Here, the number of PLS components $L$ to be estimated can be chosen by cross-validation.

@@

	\section{Notations}
	\begin{itemize}
		\item aaa
	\end{itemize}

@@

>>>>

	\subsection{Iterative steps}\label{section:sub:iterative}
	The iterative process presented here yields an orthonormal hybrid basis that effectively captures the predictor-response relationships. It proceeds through two intermediate steps: the estimation of the PLS component direction (Section \ref{section:sub:compute_PLS_component}) and residualization (Section \ref{section:sub:residualization}). 
	The proofs are deferred to Appendix \ref{section:proof:section:sub:iterative}.
	The properties of the resulting estimates are introduced in Section \ref{section:main:properties_of_our_method}.
	\subsubsection{Iterative step 1: regularized estimation of  PLS component direction}  \label{section:sub:compute_PLS_component} 
	We begin by formally introducing the core optimization problem pertinent to the PLS direction estimation, which is formulated as a generalized Rayleigh quotient (Proposition \ref{proposition:eigen_noregul}). Building upon this foundational concept, we present our regularized PLS component direction estimation step that promotes smoothness (Proposition \ref{proposition:eigen_regul}). Furthermore,  we detail an efficient computational scheme (Proposition \ref{proposition:linear_regul}).

@@

	\subsection{Final step: estimating  the hybrid regression coefficient}
	\label{section:sub:regression_coeff}
	The hybrid regression coefficient $\beta$ in model \eqref{eq: Hybrid functional model} can be written 
	as a linear combination of PLS directions:
\begin{lemma}\label{lemma:recursive}
Let use define
$
\widehat{\iota}^{[1]} := \widehat{\xi}^{[1]}.
$
For $l \ge 2$, we recursively   define:
\begin{equation*}
	\widehat{ \iota}^{[l]} = \widehat{\xi}^{[l]} - \sum_{u=1}^{l-1} \langle \widehat{ \delta}^{[u]}, \widehat{\xi}^{[l]} \rangle_{\mathbb{H}} \widehat{ \iota}^{[u]}. 
\end{equation*}
Then we have:
\begin{equation*}
	\widehat{\rho}_i^{[l]}
	=
	\langle  W_i^{[l]}, \widehat{\xi}^{[l]} \rangle_\mathbb{H}
	=
	\langle  W_i, \widehat{ \iota}^{[l]} \rangle_\mathbb{H}.
\end{equation*}
	\end{lemma}
Proof of Lemma \ref{lemma:recursive} is provided in Appendix \ref{section:proof:lemma:recursive}.

Next,  \eqref{algorithm_step:residualization_predictor} leads to the following model:
	\begin{equation*}
		Y_i = \sum \limits_{l=1}^L \widehat{\nu}^{[l]} \widehat{\rho}_i^{[l]} + \epsilon_i.
	\end{equation*}
	This model lets us to express $Y_i$ as:
	\begin{equation*}
		Y_i = \sum \limits_{l=1}^L \widehat{\nu}^{[l]} \langle W_i^{[l]}, \widehat{\xi}^{[l]} \rangle_\mathbb{H} +\epsilon_i 
		= 
		\bigl \langle W_i, \sum \limits_{l=1}^L \widehat{\nu}^{[l]} \widehat{ \iota }^{[l]}
		\bigr \rangle_\mathbb{H}+\epsilon_i,
	\end{equation*}
	which, given the uniqueness of $ \beta $, leads to
	\begin{equation*}
		\widehat{ \beta} = \sum \limits_{l=1}^L\widehat{\nu}^{[l]} \widehat{ \iota }^{[l]}.
	\end{equation*}
	Here, the number of PLS components $L$ to be estimated can be chosen by cross-validation.

@@

\begin{itemize}
	\item Appendix \ref{section:proof:lemma:hybrid_hilbert_space} proves that our defined product space is a separable Hilbert space.
	\item Appendix \ref{section:proof:section:sub:iterative}  provides the proofs for the core problem formulations and computational schemes of the iterative steps for estimating the PLS direction and performing residualization.
\end{itemize}
\section{Proof of Lemma \ref{lemma:hybrid_hilbert_space}}\label{section:proof:lemma:hybrid_hilbert_space}
	Both $\mathbb{L}^2[0,1]$ and $\mathbb{R}^p$ are Hilbert spaces, with their inner products, norms, and metrics corresponding to the terms defined in equation \eqref{eq: hybrid inner product}. A finite Cartesian product of Hilbert spaces, equipped with an $\ell_2$ norm of the component metrics (which is our metric), is also a Hilbert space.
	Furthermore, both of 	 $\mathbb{L}^2[0,1]$ and $\mathbb{R}^p$ are separable.
	For $\mathbb{L}^2[0,1]$, the set of all polynomials with rational coefficients is a countable and dense subset.
	For $\mathbb{R}^p$,  the set of all vectors with rational components forms a countable and dense subset.
A finite Cartesian product of separable spaces is also separable.
Therefore $\mathbb{H} = \bigl( \mathbb{L}^2[0,1] \bigr)^K \times \mathbb{R}^p$ is separable.
	\section{Proof of Section \ref{section:sub:iterative}}\label{section:proof:section:sub:iterative}
This section provides the proofs for the core problem formulations and computational schemes of the iterative steps for estimating the PLS direction and performing residualization.
Appendices \ref{section:proof:proposition:linear_regul_system_unregularized} and \ref{section:proof:proposition:eigen_regul} derive the eigenproblem formulations for unregularized and regularized PLS direction estimation, respectively.
Appendices \ref{section:proof:proposition:linear_regul} and \ref{section:proof:proposition:closed_form_orthgonalization} provide proofs for the simple computation schemes for regularized PLS direction estimation and residualization, respectively.	

@@

\subsection{Proof of Lemma \ref{lemma:recursive}}\label{section:proof:lemma:recursive}
